\begin{thebibliography}{9}

\bibitem{Abbott}
Abbot, Stephen.  \textit{Understanding Analysis}.  Springer-Verlag UTM.  (2010).

\bibitem{Cohn}
Cohn, Donald L..  \textit{Measure Theory}, 2nd Ed..  Springer.  (2013).

\bibitem{Coleman}
Coleman, Mark.  \textit{Continuous functions nowhere differentiable}.  \href{http://www.maths.manchester.ac.uk/~mdc/MATH20101/notesPermanant/Contsnondiff.pdf}{Constnondiff.pdf}

\bibitem{Gelbaum}
Gelbaum, Bernard R.; Olmsted, John M.~H..  \textit{Counterexamples in Analysis}.  Dover Publications Inc..  (1992).

\bibitem{GleasonMeasure}
Gleason, Jonatahn.  \textit{A homeomorphism of $\R$ that doesn't preserve lebesgue measurability}.  (2009).  \url{https://www.academia.edu/14753183/A_homeomorphism_of_mathbb_R_that_doesnt_preserve_lebesgue_measurability}.

\bibitem{GleasonHaar}
Gleason, Jonathan.  \textit{Existence and uniqueness of haar measure}.  (2010).  \url{https://www.academia.edu/740208/Existence_and_Uniqueness_of_Haar_Measure}.

\bibitem{Grayson}
Grayson, Daniel R..  \textit{Zorn's Lemma}.

\bibitem{Hart}
Hart, K.~P.; Nagata, J.; Vaughn, J.~E..  \textit{Encyclopedeia of General Topology}.  Elsevier Science.  (2004).

\bibitem{Honig}
H\"{o}nig, Chaim Samuel.  \textit{Proof of the well-ordering of cardinal numbers}

\bibitem{Howes}
Howes, Norman R..  \textit{Modern Analysis and Toology}.  Springer-Verlag Universitext.  (1991).

\bibitem{Isbell}
Isbell, J.~R.~.  \textit{Uniform spaces}.  American Mathematical Society.  (1964).

\bibitem{Kelley}
Kelley, John L..  \textit{General Topology}.  Springer-Verlag GTM.  (1955).

\bibitem{mathoverflow}
\href{http://mathoverflow.net/}{mathoverflow}

\bibitem{stackexchange}
\href{http://math.stackexchange.com/}{math.stackexchange}

\bibitem{Munkres}
Munkres, James.  \textit{Topology}, 2nd Ed..  Prentice Hall.  (2000).

\bibitem{Pugh}
Pugh, Charles Chapman.  \textit{Real Mathematical Analysis}.  Springer-Verlag UTM.  (2002).

\bibitem{Ross}
Ross, Kenneth.  \textit{Elementary Analysis:  The Theory of Calculus}.  Springer-Verlag UTM.  (2013).

\bibitem{Rudin}
Rudin, Walter.  \textit{Principles of Mathematical Analysis}, 3rd Ed..  McGraw-Hill.  (1976).

\bibitem{BigRudin}
Rudin, Walter.  \textit{Real and Complex Analysis}, 3rd Ed..  McGraw-Hill.  (1987).

\bibitem{Steen}
Steen, Lynn Arthur.  Seebach Jr., J.~Arthur.  \textit{Counterexamples in Topology}.  Holt, Rinehart and Winston Inc..  (1970).

\bibitem{Stein}
Stein, Elias M.  Shakarchi, Rami.  \textit{Real Analysis---Meas Theory, Integration, and Hilbert Spaces}.  Princeton University Press.  (2005).

\bibitem{Thomas}
Thomas, J..  \textit{A regular space, not completely regular}.  The American Mathematical Monthly, Vol.~76, No.~2, pp.~181--182.  (1969).

\bibitem{Wikipedia}
Wikipedia

\end{thebibliography}